---
title: "Practice 2"
output: html_notebook
---

## M1

Set up grid for all subquestions.

```{r}
mygrid <- seq(0, 1, .05)

```

WWW

```{r}
likelihood <- dbinom(3, 3, mygrid)
prior <- rep(1, length(mygrid))
unstd_posterior <- likelihood*prior
posterior <- unstd_posterior / sum(unstd_posterior)
plot(mygrid, posterior, type = "b")

```

WWWL

```{r}
likelihood <- dbinom(3, 4, mygrid)
prior <- rep(1, length(mygrid))
unstd_posterior <- likelihood*prior
posterior <- unstd_posterior / sum(unstd_posterior)
plot(mygrid, posterior, type = "b")

```

LWWLWWW

```{r}
likelihood <- dbinom(5, 7, mygrid)
prior <- rep(1, length(mygrid))
unstd_posterior <- likelihood*prior
posterior <- unstd_posterior / sum(unstd_posterior)
plot(mygrid, posterior, type = "b")

```

## M2

Get new prior and just do the last one from above.

```{r}
likelihood <- dbinom(5, 7, mygrid)
prior <- ifelse(mygrid < .5, 0, 1)
unstd_posterior <- likelihood * prior
posterior <- unstd_posterior / sum(unstd_posterior)
plot(mygrid, posterior, type = "b")


```

## M3

We're ~3.33 times more likely to see land if it's Mars rather than Earth. So we can use the odds version of Bayes' rule to get the posterior probability that it's Mars (and then Earth).

```{r}
3.333 / 4.333

1 - (3.333 / 4.333)

```


[...]

## H1

The LR for species B is 2 so the posterior odds are 2/1 and thus the posterior probability is 2/3. So the probability of the next birth being twins is a 1/3 mix of 10% and a 2/3 mix of 20%. So the probability next time is .167.

