---
title: "Chapter 5 (video 4)"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages.

```{r, message=FALSE}
library(tidyverse)
library(rethinking)
library(tidybayes)
library(tidybayes.rethinking)
library(ggthemes)
theme_set(theme_pander())
```

## Lecture 4

Data prep.

```{r}
data(Howell1)
d <- Howell1 |> filter(age >= 18)
d2 <- d |> 
  rename(W = weight,
         H = height,
         A = age,
         S = male) |> 
  mutate(S = as.integer(S+1))
```

Formula for estimating "causal system."

```{r}
Hbar <- mean(d2$H)
  
flist <- 
  alist(
    
  # weight model
    W ~ dnorm(mu, sigma),
    mu <- alpha[S] + beta[S]*(H-Hbar),
    alpha[S] ~ dnorm(45, 20),
    beta[S] ~ dlnorm(0, 1),
    sigma ~ dunif(0, 30),
    
  # height model
    H ~ dnorm(nu, tau),
    nu <- gamma[S],
    gamma[S] ~ dnorm(155, 30),
    tau ~ dunif(0, 30)
  
  )

fit <- quap(flist,
            data = d2)
```

Posterior summary

```{r}
precis(fit, depth = 2) # need depth = 2 when you have "vector" parameters like index variables
```

Get 10,000 draws from the posterior.

```{r}
pdraws <- tidy_draws(fit, n = 1e4) |> 
  janitor::clean_names() |>    # fixes weird backticks and brackets
  select(-chain, -iteration)   # drops MCMC things we don't need
```

Now we can use this full posterior to make inferences about parameters and parameter differences.

For example, we might ask: what is the difference in the height -> weight slope between males (sex = 2) and females (sex = 1)?

```{r}
pdraws |> 
  mutate(bdiff = beta_2 - beta_1) |>
  pull(bdiff) |> 
  HPDI()
```

This is the 89% confidence interval of the difference in kg/cm between males and females. 

Or we could estimate the total effect of sex on the height difference.

```{r}
pdraws |> 
  mutate(gdiff = gamma_2 - gamma_1) |> 
  pull(gdiff) |> 
  HPDI()
```

This is the total sex-caused difference in height in cm. We could also plot it.

```{r}
pdraws |> 
  mutate(gdiff = gamma_2 - gamma_1) |> 

ggplot(aes(x = gdiff)) +
  geom_density()
```

Now RM wants to do an intervention where he calculates the intervention effect of changing natal sex (around 46:15 on the video). His code looks monstrous to me but this is what I think I'm seeing.

We want to know what would happen if we changed "female" to "male". We need to calculate both the S -> W path and the S -> H -> W path.

We need to calculate two new simulated variables:

1. The data distribution for an average female implied by the model
2. The data distribution for an average male implied by the model

Then:

3. Take the difference between them.

We need to incorporate FIVE kinds of uncertainty in each of (1) and (2):

- uncertainty about each sex's average height (with no other variables; i.e., uncertainty about gamma)
- uncertainty about the location of a *particular height* conditional on that average (tau)
- uncertainty about each sex's average weight at the average height (alpha)
- uncertainty about the relationship between height and weight (uncertainty about beta)
- uncertainty about a *particular height* conditional on the above (sigma)

```{r}
hbar <- mean(d2$H)

pdraws <- pdraws |>
  mutate(m_height = rnorm(nrow(pdraws), gamma_2, tau),
         f_height = rnorm(nrow(pdraws), gamma_1, tau),
         m_weight = rnorm(nrow(pdraws), beta_2 * (m_height - hbar), sigma),
         f_weight = rnorm(nrow(pdraws), beta_1 * (f_height - hbar), sigma),
         do_s = m_weight - f_weight)

```

We can plot the effect altering birth sex would have on weight in kg.

```{r}
data_sim_plot <- 
  ggplot(pdraws,
         aes( x = do_s)) +
  geom_density() +
  geom_vline(xintercept = 0,
             color = "red",
             linetype = "dashed")

data_sim_plot
```

We could use these calculations to estimate what psychologists call "the probability of superiority"; that is, what percentage of the time males would be expected to be taller than females because of the "treatment."

```{r}
mean(pdraws$do_s > 0)
```

I could do all of the above getting rid of the "residual" variance. There are easier ways, but why not this?

```{r}
hbar <- mean(d2$H)

pdraws2 <- pdraws |>
  mutate(m_height = rnorm(nrow(pdraws), gamma_2, 0),
         f_height = rnorm(nrow(pdraws), gamma_1, 0),
         m_weight = rnorm(nrow(pdraws), beta_2 * (m_height - hbar), 0),
         f_weight = rnorm(nrow(pdraws), beta_1 * (f_height - hbar), 0),
         do_s = m_weight - f_weight)
```

Now plot.

```{r}
par_sim_plot <- 
ggplot(pdraws2,
       aes( x = do_s)) +
  geom_density()

par_sim_plot
```

```{r}
library(patchwork)
p1 <- data_sim_plot + scale_x_continuous(limits = c(-20, 30)) 
p2 <- par_sim_plot + scale_x_continuous(limits = c(-20, 30))

p1/p2
```


## Lecture 5

Data prep.

```{r}
data("WaffleDivorce")
d <- WaffleDivorce |> 
  select(Loc, Divorce, Marriage, MedianAgeMarriage, South) |> 
  rename(state = Loc,
         D = Divorce,
         M = Marriage,
         A = MedianAgeMarriage,
         S = South) |> 
  mutate(across(c(D,M,A),
                ~ (.x - mean(.x)) / sd(.x)))
```

Do model.

```{r}
flist <- 
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + bM*M + bA*A,
    a ~ dnorm(0, .2),
    bM ~ dnorm(0, .5),
    bA ~ dnorm(0, .5),
    sigma ~ dexp(1)
  )

fit <- quap(flist,
            data = d)
```

Look at the posterior estimates.

```{r}
precis(fit)

precis(fit) |> plot()
```

Let's try tidy.

```{r}
post <- tidy_draws(fit, n = 1e4) |> 
  select(a:sigma)

head(post)
```

Plotting prep.

```{r}
longpost <- post |> 
  pivot_longer(everything(),
               names_to = "term",
               values_to = "values")

head(longpost, 8)
```

Plotting.

```{r}
ggplot(longpost,
       aes(x = values)) +
  geom_histogram(color = "white",
                 binwidth = .05,
                 boundary = 0) +
  facet_grid(rows = vars(term),
             switch = "y") +
  geom_vline(xintercept = 0,
             color = "gray") +
  scale_y_continuous(breaks = NULL) +
  labs(y = "")
```

Another way.

```{r}
postsum <- longpost |>
  group_by(term) |> 
  summarize(mean = mean(values),
            lb = quantile(values, .055),
            ub = quantile(values, .945))
postsum

ggplot(postsum,
       aes(y = reorder(term, c(4,3,2,1)),
           x = mean,
           xmin = lb,
           xmax = ub)) +
  geom_point(size = 2) +
  geom_linerange(size = 1) +
  geom_vline(xintercept = 0,
             color = "gray") +
  xlim(-1,1) +
  labs(x = "Estimate",
       y = "")
```

## Additional stuff

```{r}
curve(dexp(x), from = 0, to = 2)
```

